---
title: "Comparing ZooMSS and ZooMizer"
author: "Patrick Sykes"
output:
  html_document:
    df_print: paged
place: Brisbane
---

# Introduction

Here we are trying to compare the original code ZooMSS model (version 2, Heneghan et al. 2020) with ZooMizer, an attempt at recreating it within the mizer package.

We start by initialising and running the two models.


```{r setup}
library(mizer)
library(Rcpp)

source("fZooMSS_Model.R") #source the model code
source("fZooMSS_Setup.R")
source("fZooMSS_CalculatePhytoParam.R")
#Rcpp::sourceCpp("fZooMSS_MvF_Rcpp.cpp") #explicitly using same numerics as mizer
#setwd("~/GitHub/ZooMizer/")
## Build environmental data
#enviro <- readRDS("data/enviro_test20.RDS")
enviro <- readRDS("data/envirofull_20200317.RDS")
enviro_data <- fZooMSS_CalculatePhytoParam(enviro) # Calculate Phytoplankton Parameters
Groups <- read.csv("TestGroups.csv") #[c(1,3,12),]
#Groups$Carbon <- 0.1
#Groups$Prop[-3] <- 0



dt <- 001
input_params <- enviro_data[1,]
input_params$dt <- dt
input_params$tmax <- 2
SaveTimeSteps <- TRUE

zoomsstest <- list()
zoomsstest$model$model_runtime <- system.time(
 zoomsstest <- fZooMSS_Model(input_params, Groups, SaveTimeSteps)
)
saveRDS(zoomsstest,"zoomssoutputdt1.rds")
zoomsstest <- readRDS("zoomssoutputdt1.rds")

input_params$tmaxx <- input_params$tmax

source("fZooMizer_run.R")
environment(new_project_simple) <- asNamespace('mizer')
assignInNamespace("project_simple", new_project_simple, ns = "mizer")

environment(new_newMultispeciesParams) <- asNamespace('mizer')
assignInNamespace("newMultispeciesParams", new_newMultispeciesParams, ns = "mizer")

environment(new_emptyParams) <- asNamespace('mizer')
assignInNamespace("emptyParams", new_emptyParams, ns = "mizer")

#setValidity("MizerParams", new_validMizerParams)

groups <- read.csv("data/TestGroups_mizer.csv") #[c(1,3),]
#groups$Carbon <- 0.1
#groups$Prop[-3] <- 0

zoomizertest <- fZooMizer_run(groups = groups, input = input_params)
```

Let's try and look at some comparisons...

# Initial values

We want to check that the initial setups were all the same. the `all.equal` function tests the 'near equality' of two objects, that is, whether the objects are equal to within a tolerance (by default the tolerance is 1.5e-8). It's appropriate to use that here to allow for small floating point errors introduced by unit conversions and converting back and forth between absolute and logged values.

```{r collapse=TRUE}

#weight vectors (i.e. size classes)
length(zoomsstest$model$param$w_phyto) == sum(zoomizertest@n_pp[1,] > 0) #phyto cuts off at same place?

all.equal(zoomsstest$model$param$w, zoomizertest@params@w, check.attributes=FALSE)

all.equal(zoomsstest$model$param$w_phyto, zoomizertest@params@w_full[1:length(zoomsstest$model$param$w_phyto)],check.attributes=FALSE)

#abundances - phyto - dimensions time x weight
all.equal(zoomizertest@n_pp[1,1:length(zoomsstest$model$param$w_phyto)] *
            zoomizertest@params@dw_full[1:length(zoomsstest$model$nPP)],
          zoomsstest$model$nPP,check.attributes=FALSE)

#abundances - zoo and fish. Note N is in dimensions time x weight x species
all.equal(zoomsstest$model$N[1,,] , zoomizertest@params@initial_n*matrix(zoomizertest@params@dw,nrow=zoomsstest$model$param$ngrps, ncol = length(zoomsstest$model$param$w), byrow=TRUE), check.attributes=FALSE)
#View(zoomsstest$model$N[1,,] / matrix(zoomsstest$model$param$w,nrow=zoomsstest$model$param$ngrps, ncol = length(zoomsstest$model$param$w), byrow=TRUE))
#View(zoomizertest@params@initial_n)

#initialnsaved <- readRDS("data/initialn.RDS")
all.equal(zoomsstest$model$N[1,,] , sweep(zoomizertest@params@initial_n, 2, zoomizertest@params@dw, "*"), check.attributes=FALSE)
# quick check here: abundances after a time step?
all.equal(sweep(zoomizertest@n[2,,],2, zoomizertest@params@dw, "*"), zoomsstest$model$N[2,,] , check.attributes=FALSE)


```

So we have that the size classes and initial abundance values (taking into account that ZooMSS is in absolute abundance and ZooMizer is in normalised abundance) are all in agreement. However, after a single time step they differ by a lot! Let's look into what might be doing this.

Recall the McKendrick-von Foerster equation

$$ \frac{\partial}{\partial t} N_i(w,t) = - \frac{\partial}{\partial w} (g_i(w,t) N_i(w,t)) - \mu_i(w,t)N_i(w,t), $$
where $g_i$ is growth (which is due to feeding) and $\mu_i$ is mortality (due to senescence and predation).

Let's look at the first term on the right-hand side: growth.

# Growth due to feeding

Trying with the actual growth matrices. Both Mizer and ZooMSS use g/yr as the units.

```{r}

param <- zoomsstest$model$param
model <- zoomsstest$model

growth_multiplier <- colSums(model$N[1,,] * model$assim_eff) # 1 x n_sizes
predation_multiplier <- model$N[1,,] * model$temp_eff # n_species x n_sizes

### DO GROWTH
dynam_growthkernel <- model$dynam_growthkernel
dim(dynam_growthkernel) <- c(param$ngrps*param$ngrid, param$ngrid)
cs <- .colSums(growth_multiplier * t(dynam_growthkernel), m = param$ngrid, n = param$ngrps*param$ngrid)
dim(cs) <- c(param$ngrps, param$ngrid)
gg <- model$ingested_phyto + cs
  
zoomizererepro <- getERepro(zoomizertest@params) #all 0 as expected
zoomizerereprogrowth <- getEReproAndGrowth(zoomizertest@params)
zoomizeregrowth <- getEGrowth(zoomizertest@params)


all.equal(zoomizeregrowth, zoomizerereprogrowth) #should be equal since no energy goes to repro

all.equal(zoomizeregrowth[,2:178], gg[,1:177], check.attributes=FALSE) #it was really close earlier with zoomizeregrowth/w, now mizer's N is in 1/g so perhaps this is the right conversion?
```

Looking at the matrices, it's clearly not right. Though it's not completely clear whether the difference abundance units may be affecting this. Consider the function for growth:

$$ g_i(w,t) = \tau \sum_{j} 2.5 C_j V_i(w)\int_{w_p}^{w} \phi_i (w,w')N_j(w',t)w'dw' $$

The growth in ZooMSS is calculated from temperature effect constant $\tau$, prey carbon content $C_j$, the search volume $V_i(w)$, and prey size selectivity function $\phi_i(w,w')$. Let's look at these one at a time:

## Temperature effect

```{r}

all.equal(zoomsstest$model$temp_eff, zoomizertest@params@other_params$temp_eff)

```

## Prey carbon / assimilation efficiency

Note that in ZooMSS, the assimilation efficiency is a (prey species) $\times$ (prey size) matrix, while it's a (predator species) $\times$ (prey species) matrix in Mizer. This is simply fixed for each prey species though, so we can compare the ZooMSS value with the transpose of the ZooMizer value.

```{r}

all.equal(zoomsstest$model$assim_eff[,1:12], t(zoomizertest@params@other_params$assim_eff))

```


## Search volume

```{r}

#zoomss code
SearchVol <- matrix(NA, nrow = zoomsstest$model$param$ngrps, ncol = zoomsstest$model$param$ngrid)
for(i in 1:zoomsstest$model$param$ngrps){
        ### Search volume
    SearchVol[i,] <- (zoomsstest$model$param$Groups$SearchCoef[i])*(zoomsstest$model$param$w^(zoomsstest$model$param$Groups$SearchExp[i]))
    SearchVol[i, 10^(zoomsstest$model$param$Groups$Wmax[i]) < zoomsstest$model$param$w] <- 0
    SearchVol[i, 10^(zoomsstest$model$param$Groups$W0[i]) > zoomsstest$model$param$w] <- 0
    
}


all.equal(SearchVol, zoomizertest@params@search_vol, check.attributes=FALSE)

```

These are in agreement. How about feeding kernels?

```{r}

  
  gg_log_t_phyto <- ((param$w^-1) %*% t(param$w_phyto))/log(10) # Growth
  diet_log_t_phyto <- matrix(param$w_phyto, nrow = length(param$w), ncol = length(param$w_phyto), byrow = TRUE) # Diet/Ingestion

  # Predators are rows, dynam prey weights are columns
  gg_log_t_dynam <- ((param$w^-1) %*% t(param$w))/log(10) # Growth
  diet_log_t_dynam <- matrix(param$w, nrow = length(param$w), ncol = length(param$w), byrow = TRUE) # Diet/ingestion

  ### PREDATION KERNELS FOR PHYTOPLANKTON SPECTRUM AND DYNAMIC SPECTRUM
  phyto_pred_weight_matrix <- matrix(param$w, nrow = param$ngrid, ncol = param$ngridPP)
  dynam_pred_weight_matrix <- matrix(param$w, nrow = param$ngrid, ncol = param$ngrid)
  phyto_prey_weight_matrix <- matrix(param$w_phyto, nrow = param$ngrid, ncol = param$ngridPP, byrow = TRUE)
  dynam_prey_weight_matrix <- matrix(param$w, nrow = param$ngrid, ncol = param$ngrid, byrow = TRUE)

  ## Search Volume storage
  SearchVol <- matrix(NA, nrow = param$ngrps, ncol = param$ngrid) # Search volume

  # Simpson's Rule matrices for growth, diffusion and mortality integrals
  simp_phyto <- array(1, dim = param$ngridPP)
  simp_phyto[c(seq(2, param$ngridPP-1,2))] <- 4
  simp_phyto[c(seq(3, param$ngridPP-1,2))] <- 2
  sm_phyto <- matrix(simp_phyto, nrow = param$ngrid, ncol = param$ngridPP, byrow = TRUE) * (param$dx/3)

  simp_dynam <- array(1, dim = param$ngrid)
  simp_dynam[c(seq(2, param$ngrid-1,2))] <- 4
  simp_dynam[c(seq(3, param$ngrid-1,2))] <- 2
  sm_dynam <- matrix(simp_dynam, nrow = param$ngrid, ncol = param$ngrid, byrow = TRUE) * (param$dx/3)
  
   ### Q10 OF 2 FOR ALL ZOO AND FISH
  temp_zoo <- rep(2.^((param$sst - 30)/10), param$num_zoo) # exp(23.93 - 0.59/(8.62e-05*(273+environ$sst)))
  temp_fish <- rep(2.^((param$sst - 30)/10), param$num_fish)
  temp_eff <- matrix(c(temp_zoo, temp_fish), nrow = param$ngrps, ncol = param$ngrid)

  
  phyto_predkernel <- array(0, dim=c(param$ngrps,param$ngrid,param$ngridPP))
  dynam_predkernel <- array(0, dim=c(param$ngrps,param$ngrid,param$ngrid))
  #### CALCULATES CONSTANT BITS OF THE MODEL FUNCTIONS FOR EACH GROUP
  for(i in 1:param$ngrps){
    ## Senescence mortality
    if(i < 10){
      model$M_sb[i,] <- param$ZSpre*(param$w/(10^(param$Groups$Wmat[i])))^param$ZSexp
      model$M_sb[i, 10^(param$Groups$Wmax[i]) < param$w] <- 0
      model$M_sb[i, 10^(param$Groups$Wmat[i]) > param$w] <- 0
    }

    if(i > 9){
      model$M_sb[i,] <- 0.1*param$ZSpre*(param$w/(10^(param$Groups$Wmat[i])))^param$ZSexp
      model$M_sb[i, 10^(param$Groups$Wmax[i]) < param$w] <- 0
      model$M_sb[i, 10^(param$Groups$Wmat[i]) > param$w] <- 0
    }

    ### Search volume
    SearchVol[i,] <- (param$Groups$SearchCoef[i])*(param$w^(param$Groups$SearchExp[i]))
    SearchVol[i, 10^(zoomsstest$model$param$Groups$Wmax[i]) < zoomsstest$model$param$w * (1 + 1e-06)] <- 0
    SearchVol[i, 10^(zoomsstest$model$param$Groups$W0[i]) > zoomsstest$model$param$w * (1 - 1e-06)] <- 0

    ### Predation Kernels
    if(is.na(param$Groups$PPMRscale[i]) == FALSE){ # If group has an m-value (zooplankton)
      # Calculate PPMR for zooplankton, which changes according to body-size (Wirtz, 2012)
      D.z <- 2*(3*param$w*1e12/(4*pi))^(1/3) # convert body mass g to ESD (um)
      betas <- (exp(0.02*log(D.z)^2 - param$Groups$PPMRscale[i] + 1.832))^3 # Wirtz's equation
      beta_mat_phyto <- matrix(betas, nrow = param$ngrid, ncol = param$ngridPP)
      beta_mat_dynam <- matrix(betas, nrow = param$ngrid, ncol = param$ngrid)

      # Calculate feeding kernels
      phyto_predkernel[i,,] <- exp(-0.5*(log((beta_mat_phyto*phyto_prey_weight_matrix)/
                                            phyto_pred_weight_matrix)/param$Groups$FeedWidth[i])^2)/
        sqrt(2*pi*param$Groups$FeedWidth[i]^2)
      dynam_predkernel[i,,] <- exp(-0.5*(log((beta_mat_dynam*dynam_prey_weight_matrix)/
                                            dynam_pred_weight_matrix)/param$Groups$FeedWidth[i])^2)/
        sqrt(2*pi*param$Groups$FeedWidth[i]^2)

      # The feeding kernal of filter feeders is not expected to change much with increasing size so we fix it here

      # if (param$fixed_filterPPMR == TRUE){
        if(i == 3){
          phyto_predkernel[i,,] <- matrix(phyto_predkernel[i,44,], nrow = param$ngrid, ncol = param$ngridPP, byrow = TRUE)
          dynam_predkernel[i,,] <- matrix(dynam_predkernel[i,44,], nrow = param$ngrid, ncol = param$ngrid, byrow = TRUE)
        }
        if(i == 8){
          phyto_predkernel[i,,] <- matrix(phyto_predkernel[i,61,], nrow = param$ngrid, ncol = param$ngridPP, byrow = TRUE)
          dynam_predkernel[i,,] <- matrix(dynam_predkernel[i,61,], nrow = param$ngrid, ncol = param$ngrid, byrow = TRUE)
        }
      # }

    } else { # If group does not have an m-value (fish)
      beta_mat_phyto <- matrix(param$Groups$PPMR[i], nrow = param$ngrid, ncol = param$ngridPP)
      beta_mat_dynam <- matrix(param$Groups$PPMR[i], nrow = param$ngrid, ncol = param$ngrid)

      # Calculate feeding kernels
      phyto_predkernel[i,,] <- exp(-0.5*(log((beta_mat_phyto*phyto_prey_weight_matrix)/
                                            phyto_pred_weight_matrix)/param$Groups$FeedWidth[i])^2)/
        sqrt(2*pi*param$Groups$FeedWidth[i]^2)
      dynam_predkernel[i,,] <- exp(-0.5*(log((beta_mat_dynam*dynam_prey_weight_matrix)/
                                            dynam_pred_weight_matrix)/param$Groups$FeedWidth[i])^2)/
        sqrt(2*pi*param$Groups$FeedWidth[i]^2)
      
    }
    
}

  zoomizerpredkernel <- getPredKernel(zoomizertest@params)
  dim(zoomizerpredkernel)
#all.equal(sp_dynam_predkernel[7,,],zoomizerpredkernel[7,,(216-178+1):216], check.attributes=FALSE) # different dimensions - how to compare?
  
wmax_phyto <- dim(phyto_predkernel)[3]
dim(zoomizerpredkernel)

#sanity check: zoomss phyto and dynamic spectrum feeding kernels all equal:
all.equal(phyto_predkernel[,,39:wmax_phyto],dynam_predkernel[,,1:(wmax_phyto-39+1)])

#comparing phytoplankton predation kernels
all.equal(phyto_predkernel[,,], zoomizerpredkernel[,,1:wmax_phyto], check.attributes = FALSE) #TRUE  

all.equal(phyto_predkernel[1,,], zoomizerpredkernel[1,,1:wmax_phyto], check.attributes=FALSE)   #TRUE  Flagellates 
all.equal(phyto_predkernel[2,,], zoomizerpredkernel[2,,1:wmax_phyto], check.attributes=FALSE)   #TRUE  Ciliates    
all.equal(phyto_predkernel[3,,], zoomizerpredkernel[3,,1:wmax_phyto], check.attributes=FALSE)   #TRUE  Larvaceans  
all.equal(phyto_predkernel[4,,], zoomizerpredkernel[4,,1:wmax_phyto], check.attributes=FALSE)   #TRUE  OmniCopepods
all.equal(phyto_predkernel[5,,], zoomizerpredkernel[5,,1:wmax_phyto], check.attributes=FALSE)   #TRUE  CarnCopepods
all.equal(phyto_predkernel[6,,], zoomizerpredkernel[6,,1:wmax_phyto], check.attributes=FALSE)   #TRUE  Euphausiids
all.equal(phyto_predkernel[7,,], zoomizerpredkernel[7,,1:wmax_phyto], check.attributes=FALSE)   #TRUE  Chaetognaths
all.equal(phyto_predkernel[8,,], zoomizerpredkernel[8,,1:wmax_phyto], check.attributes=FALSE)   #TRUE  Salps
all.equal(phyto_predkernel[9,,], zoomizerpredkernel[9,,1:wmax_phyto], check.attributes=FALSE)   #TRUE  Jellyfish
all.equal(phyto_predkernel[10,,], zoomizerpredkernel[10,,1:wmax_phyto], check.attributes=FALSE) #TRUE  Fish_Small
all.equal(phyto_predkernel[11,,], zoomizerpredkernel[11,,1:wmax_phyto], check.attributes=FALSE) #TRUE  Fish_Med
all.equal(phyto_predkernel[12,,], zoomizerpredkernel[12,,1:wmax_phyto], check.attributes=FALSE) #TRUE  Fish_Large

#comparing zoo/fish predation kernels
dim(dynam_predkernel)
all.equal(dynam_predkernel[,,], zoomizerpredkernel[,,39:216], check.attributes = FALSE)

all.equal(dynam_predkernel[1,,], zoomizerpredkernel[1,,39:216], check.attributes=FALSE)   #TRUE  Flagellates 
all.equal(dynam_predkernel[2,,], zoomizerpredkernel[2,,39:216], check.attributes=FALSE)   #TRUE  Ciliates
all.equal(dynam_predkernel[3,,], zoomizerpredkernel[3,,39:216], check.attributes=FALSE)   #TRUE  Larvaceans  
all.equal(dynam_predkernel[4,,], zoomizerpredkernel[4,,39:216], check.attributes=FALSE)   #TRUE  OmniCopepods
all.equal(dynam_predkernel[5,,], zoomizerpredkernel[5,,39:216], check.attributes=FALSE)   #TRUE  CarnCopepods
all.equal(dynam_predkernel[6,,], zoomizerpredkernel[6,,39:216], check.attributes=FALSE)   #TRUE  Euphausiids
all.equal(dynam_predkernel[7,,], zoomizerpredkernel[7,,39:216], check.attributes=FALSE)   #TRUE  Chaetognaths
all.equal(dynam_predkernel[8,,], zoomizerpredkernel[8,,39:216], check.attributes=FALSE)   #TRUE  Salps
all.equal(dynam_predkernel[9,,], zoomizerpredkernel[9,,39:216], check.attributes=FALSE)   #TRUE  Jellyfish
all.equal(dynam_predkernel[10,,], zoomizerpredkernel[10,,39:216], check.attributes=FALSE) #TRUE  Fish_Small
all.equal(dynam_predkernel[11,,], zoomizerpredkernel[11,,39:216], check.attributes=FALSE) #TRUE  Fish_Med
all.equal(dynam_predkernel[12,,], zoomizerpredkernel[12,,39:216], check.attributes=FALSE) #TRUE  Fish_Large


```

Ingested prey:

```{r}
zoomizerdiet <- getDiet(zoomizertest@params, proportion = FALSE) 
  #zoomssgrid[[1]]
dim(zoomizerdiet)

# dimnames(zoomizerdiet)
# zoomizerdiet[,,13] #rowSums(sweep(params@pred_kernel, 3, params@dw_full * params@w_full * n_pp, "*"), dims = 2)
dim(zoomsstest$model$phyto_dietkernel)
  
all.equal(zoomsstest$model$phyto_dietkernel[1,2,1],zoomsstest$model$phyto_dietkernel[1,2,2]) #sanity check

#phyto ingestion
#model$temp_eff*(rowSums(sweep(model$phyto_growthkernel, 3, model$nPP, "*"), dims = 2))
all.equal(zoomizerdiet[,,13], zoomsstest$model$ingested_phyto, check.attributes = FALSE)

```


## Growth integrals

It looks like the difference may be in the different methods being used: ZooMSS does its numerical integration over log(weight) unlike Mizer, which does it over absolute weight. The conversion for this is given by arrays `dynam_growthkernel` and `phyto_growthkernel` and related helper arrays `gg_log_t_dynam` etc. These are what I will look at next...

Here's the setup in ZooMSS:

```{r}

param <- zoomsstest$model$param
model <- zoomsstest$model

ngrps <- zoomsstest$model$param$ngrps
ngrid <- zoomsstest$model$param$ngrid

idx_iter <- 2:ngrid
idx <- 2:(ngrid-1)
  
  # Matrices for MvF and MvF-D numeric solution
  A_iter <- matrix(0, nrow = ngrps, ncol = ngrid)
  C_iter <- matrix(0, nrow = ngrps, ncol = ngrid)
  S_iter <- matrix(0, nrow = ngrps, ncol = ngrid)

  A <- matrix(0, nrow = ngrps, ncol = ngrid)
  B <- matrix(0, nrow = ngrps, ncol = ngrid)
  C <- matrix(0, nrow = ngrps, ncol = ngrid)
  S <- matrix(0, nrow = ngrps, ncol = ngrid)

  # Temporary Matrices that get updated each time step some of these saved for output
  N <- matrix(model$N[1,,], nrow = ngrps, ncol = ngrid) # Abundances of functional groups, dim 1 = groups, dim 2 = size classes

    growth_multiplier <- colSums(N * model$assim_eff) # 1 x n_sizes
    predation_multiplier <- N * model$temp_eff # n_species x n_sizes
    diffusion_multiplier <- colSums(N * (model$assim_eff^2)) # 1 x n_sizes

    ### DO GROWTH
    dim(model$dynam_growthkernel) <- c(ngrps*ngrid, ngrid)
    cs <- .colSums(growth_multiplier * t(model$dynam_growthkernel), m = ngrid, n = ngrps*ngrid)
    dim(cs) <- c(ngrps, ngrid)
    gg <- model$ingested_phyto + cs
    gg_dynam <- cs

    sw2 <- sweep(model$dynam_mortkernel, c(2,3), predation_multiplier, '*') # n_sizes x n_species x n_sizes
    ap2 <- aperm(sw2, c(2,3,1))
    M2 <- .colSums(colSums(ap2),ngrid,ngrid) # 1 x n_sizes
    Z <- sweep(model$M_sb + model$fish_mort, 2, M2, '+') # Total dynamic spectrum mortality (n_species x n_sizes)
    rm(sw2, ap2)

    ### DO DIFFUSION
    dim(model$dynam_diffkernel) <- c(ngrps*ngrid, ngrid)
    cs <- .colSums(diffusion_multiplier * t(model$dynam_diffkernel), m = ngrid, n = ngrps*ngrid)
    dim(cs) <- c(ngrps, ngrid)
    diff <- model$diff_phyto + cs

    # sw3 <- sweep(dynam_diffkernel, 3, diffusion_multiplier, '*')
    # ap3 <- aperm(sw3, c(3,1,2))
    # cs3 <- colSums(ap3)
    # diff <- diff_phyto + cs3
    # rm(sw3, ap3, cs3)

    ### MvF WITH DIFFUSION ALGORITHM
    # Numerical implementation matrices (for MvF without diffusion)
    A_iter[,idx_iter] <- param$dt/param$dx * gg[,idx_iter-1] # Growth stuff
    C_iter[,idx_iter] <- 1 + param$dt * Z[,idx_iter] + param$dt/param$dx * gg[,idx_iter] # Mortality
    S_iter[,idx_iter] <- N[,idx_iter] # N at.....
    N_iter <- N # Current Abundance
    N_iter[1,1] <- N[1,1] # This forces R to make a copy of the variable. Otherwise N is linked to N_iter in the Rcpp code and they change together.

    # # Numerical implementation matrices (for MvF WITH diffusion)
    # A[,idx] <- param$dt/param$dx * (gg[,idx-1] + diff[,idx-1] * (log(10)/2+1/(2*param$dx))) # Growth stuff
    # B[,idx] <- diff[,idx+1] * param$dt/(2*param$dx^2) # Diffusion term
    # C[,idx] <- 1 + param$dt * Z[,idx] + param$dt/dx*(gg[,idx] + diff[,idx] * (log(10)/2+1/dx)) # Mortality
    # S[,idx] <- N[,idx]
    # 

```


Now, the equivalent in ZooMizer:

```{r}

params <- zoomizertest@params

#dt <- 0.01

n <- params@initial_n

no_sp <- nrow(params@species_params) # number of species
  no_w <- length(params@w) # number of fish size bins
  idx <- 2:no_w
  w_max_idx <- params@w_min_idx
  for (i in 1:length(w_max_idx)) {
    w_max_idx[i] <- which(round(log10(params@w),2)==round(log10(params@species_params$w_inf[i]),2))
  }

  # Hacky shortcut to access the correct element of a 2D array using 1D notation
  # This references the egg size bracket for all species, so for example
  # n[w_min_idx_array_ref] = n[,w_min_idx]
  w_min_idx_array_ref <- (params@w_min_idx - 1) * no_sp + (1:no_sp)
  # Matrices for solver
  a <- matrix(0, nrow = no_sp, ncol = no_w)
  b <- matrix(0, nrow = no_sp, ncol = no_w)
  S <- matrix(0, nrow = no_sp, ncol = no_w)

  encounter <- getEncounter(params)
  
  
  e_growth <- new_EReproAndGrowth(params = params, n = params@initial_n, n_pp = params@initial_n_pp, n_other = params@initial_n_other, t = 0, encounter = getEncounter(params), feeding_level = getFeedingLevel(params))
  mort <- getMort(params)
    n_pp <- params@initial_n_pp
    # Iterate species one time step forward:
    # a_{ij} = - g_i(w_{j-1}) / dw_j dt
    a[, idx] <- sweep(e_growth[, idx - 1, drop = FALSE] * dt, 2,
                      params@w[idx-1], "/") * 10^(-0.1) / log(10)
    # b_{ij} = 1 + g_i(w_j) / dw_j dt + \mu_i(w_j) dt
    b[] <- 1 + sweep(e_growth * dt , 2, params@w, "/") /log(10) + r$mort * dt * 0.1
    # S_{ij} <- N_i(w_j)
    S[,idx] <- n[, idx, drop = FALSE]

    
all.equal(e_growth, gg, check.attributes = FALSE)
all.equal(sweep(a, 2, params@dw, "*"), A_iter, check.attributes = FALSE)
all.equal(C_iter, b, check.attributes = FALSE)

#just looking at the contribution from phyto ingestion:
params2 <- params
params2@initial_n <- params2@initial_n*0
e_growth_phyto <- new_EReproAndGrowth(params = params, n = params@initial_n*0, n_pp = params@initial_n_pp, n_other = params@initial_n_other *0, t = 0, encounter = getEncounter(params), feeding_level = getFeedingLevel(params))
a_phyto <- matrix(0, nrow = no_sp, ncol = no_w)
a_phyto[, idx] <- sweep(e_growth_phyto[, idx - 1, drop = FALSE] * dt / 0.1, 2,
                      params@dw[idx], "/")
all.equal(sweep(a_phyto, 2, params@dw, "*")[, idx], param$dt/param$dx * model$ingested_phyto[, idx-1])

# just the ingestion of zoo/fish:
a_dynam <- a-a_phyto
A_iter_dynam <- matrix(0, nrow = ngrps, ncol = ngrid)
A_iter_dynam[,idx] <- param$dt/param$dx * gg_dynam[, idx-1]
all.equal(a_dynam, A_iter_dynam, check.attributes = FALSE)

```

If you look at the relevant comparisons (`a` against `A` and `b` against `C`), it's apparent that the Mizer values are about 7 times too high. This might mean that something has been counted twice in the ZooMizer code?

Note that `gg` in ZooMSS is mostly made up of the ingested phyto for small organisms. What is ingested phyto for ZooMizer?

```{r}

#ZooMSS


#Zoomizer
phi_prey_background <- params@other_params$assim_phyto * params@species_params$interaction_resource *
    rowSums(sweep(
      params@pred_kernel, 3, params@dw_full * params@w_full * n_pp,
      "*", check.margin = FALSE), dims = 2)
phyto_encounter <- params@other_params$temp_eff * params@search_vol * (phi_prey_background)

#plug zoomss into zoomizer
zoomssparams <- params
zoomssparams@initial_n_pp[1:wmax_phyto] <- zoomsstest$model$nPP / params@dw_full[1:wmax_phyto]
zoomssparams@initial_n[] <- sweep(zoomsstest$model$N[1,,], 2, params@dw, "/")
#new_Encounter(params = zoomssparams, n = zoomssparams@initial_n, n_pp = zoomssparams@initial_n_pp, n_other = zoomssparams@initial_n_other, t=0)

e_growth2 <- new_EReproAndGrowth(params = zoomssparams, n = zoomssparams@initial_n, n_pp = zoomssparams@initial_n_pp, n_other = zoomssparams@initial_n_other, t = 0, encounter = getEncounter(zoomssparams), feeding_level = getFeedingLevel(zoomssparams))

all.equal(e_growth2, e_growth) #TRUE - so e_growth (mizer) and gg (ZooMSS) are being calculated differently...
#all.equal(sweep(e_growth, 2, params@dw, "/"), gg/0.1, check.attributes = FALSE)

#trying to recreate mizer encounter in ZooMSS
all.equal(zoomsstest$model$phyto_growthkernel2, zoomizertest@params@pred_kernel[,,1:wmax_phyto], check.attributes = FALSE)


all.equal(zoomsstest$model$phi_prey_background, phi_prey_background, check.attributes = FALSE)
all.equal(phyto_encounter,zoomsstest$model$phyto_encounter, check.attributes = FALSE)

```


# Mortality

Do the mortality rates agree?

```{r}


#senescence alone:
# View(zoomsstest$model$M_sb)
# View(zoomizertest@params@mu_b)
all.equal(getExtMort(params)/10, zoomsstest$model$M_sb, check.attributes = FALSE) # 10 times higher - covered in the numerical step

#predation mortality
all.equal(getPredMort(zoomizertest@params)[1,]/10, M2, check.attributes = FALSE)

#overall mortality
all.equal(getMort(params)/10, Z, check.attributes = FALSE)

```

Run ZooMSS on Zoomizer initial values:

```{r eval=FALSE, include=FALSE}
curr_min_size <- vector()
  curr_max_size <- vector()
  for (i in 1:ngrps){
    curr_min_size[i] <- which(round(log10(model$param$w), digits = 2) == param$Groups$W0[i])
    curr_max_size[i] <- which(round(log10(model$param$w), digits = 2) == param$Groups$Wmax[i])
  }

idx_iter <- 2:ngrid
  idx <- 2:(ngrid-1)
A_iter[,idx_iter] <- dt/model$param$dx * getEGrowth(params)[,idx_iter-1] # Growth stuff
    C_iter[,idx_iter] <- 1 + dt * getMort(params)[,idx_iter] + dt/model$param$dx * getEGrowth(params)[,idx_iter] # Mortality
    S_iter[,idx_iter] <- (params@initial_n*params@dw)[,idx_iter] # N at.....
    N_iter <- (params@initial_n*params@dw) # Current Abundance
    N_iter[1,1] <- (params@initial_n*params@dw)[1,1] # This forces R to make a copy of the variable. Otherwise N is linked to N_iter in the Rcpp code and they change together.

    # Numerical implementation matrices (for MvF WITH diffusion)
    A[,idx] <- dt/model$param$dx * (getEGrowth(params)[,idx-1] + diff[,idx-1] * (log(10)/2+1/(2*model$param$dx))*0) # Growth stuff
    B[,idx] <- diff[,idx+1] * dt/(2*model$param$dx^2)*0 # Diffusion term
    C[,idx] <- 1 + dt * getMort(params)[,idx] + dt/model$param$dx*(getEGrowth(params)[,idx] + diff[,idx] * (log(10)/2+1/model$param$dx)*0) # Mortality
    S[,idx] <- (params@initial_n*params@dw)[,idx]

    # The original Base R code for the MvF equation
    N2 <- fZooMSS_MvF_BaseR(ngrps, curr_min_size, curr_max_size,
                            A_iter, C_iter, N_iter, S_iter,
                            A, B, C, N, S)
    
View(N2 / model$N[2,,])
    
```
