---
title: "How To ZooMizer"
author: "Patrick Sykes"
date: "07/06/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this document, I will describe the ZooMizer model and show how it runs.

In several places in the code I have included lines that are commented out: these will give you an idea of different ways that the model may be used and experiments that I have run in the past.

First, load required packages:

```{r packages}
library(mizer)
library(tidyverse)
library(assertthat)
library(foreach)
```

## Set up the environmental data

Give the job a name and set up the environment data:

```{r}

jobname <- "20220706_HowToZooMizer_15deg_chlo1"

# enviro <- data.frame(sst = 15, chlo = 10^seq(-1.5, 1.5, by=0.125))
enviro <- data.frame(sst = 15, chlo = 1)
source("fZooMSS_CalculatePhytoParam.R")
enviro <- fZooMSS_CalculatePhytoParam(enviro)
enviro$tmax <- 10
enviro$dt <- 0.01

```

The line commented out there is an example of how I would usually run over several cells; I would then usually set up a virtual cluster and use the `foreach` and `doParallel` packages as so:

```{r cluster, eval=FALSE}
library(doParallel)
cl <- makePSOCKcluster(max(1, detectCores()-1), outfile="")     ## set up cores-1 machines
registerDoParallel(cl, cores = (max(1, detectCores()-1)))
clusterEvalQ(cl, lapply(c("mizer", "assertthat", "tidyverse"), require, character.only = TRUE)) %>% invisible()
```

The following code would then be wrapped in a function and run with `foreach(ID=1:25) %dopar%`.

## Run ZooMSS to get initial conditions

The initial conditions that are used to initialise and parametrise ZooMizer are drawn from the (averaged) output of a ZooMSS run (that is run under ZooMizer) as so:

```{r ZooMSS, eval=FALSE}

Groups <- read.csv("data/TestGroups_mizer.csv") # Load in functional group information

zoomssjobname <- "20220607_HowToZooMizer_15deg_chlo1" #job name used on queue

# load in required functions
source("uncoupledmodel.R")
source("ZooMizerResourceFunctions.R")

# switch Mizer to using ZooMSS numerics
environment(new_project_simple) <- asNamespace('mizer')
assignInNamespace("project_simple", new_project_simple, ns = "mizer")

# Run the ZooMSS model inside Mizer:
phyto_fixed <- function(params, n, n_pp, n_other, rates, dt, kappa = 10^enviro$phyto_int[i], lambda= 1-enviro$phyto_slope[i], ... ) {
  npp <- kappa*params@w_full^(-lambda) #returns the fixed spectrum at every time step
  npp[params@w_full > params@resource_params$w_pp_cutoff] <- 0
  return(npp)
}

zoomss <- fZooMizer_run(groups = Groups, input = enviro, no_w = (1*177+1)) # dx=0.1
saveRDS(zoomss, file = paste0("Output/", zoomssjobname, "_ZooMSS.RDS"))

```

## Set up ZooMizer

First, we'll process the environmental data and output from the ZooMSS run above

```{r zoomizerenviro}

source("uncoupledmodel.R")
source("ZooMizerResourceFunctions.R")

# revert to Mizer standard numerics (for the fish model - zooplankton numerics still follow ZooMSS)
environment(project_simple) <- asNamespace('mizer')
assignInNamespace("project_simple", project_simple, ns = "mizer")

groups <- read.csv("data/TestGroups_mizer.csv") #load in ZooMSS groups
  
#lower "starting proportion" of salps and chaetognaths
# groups$Prop[c(7,8)] <- 0.1*groups$Prop[c(7,8)]
# #lower "starting proportion" of euphausiids
# groups$Prop[6] <- 0.1*groups$Prop[c(6)]

# change start of senescence for salps and euphs
# groups$w_mat[c(6,8)] <- groups$w_mat[c(6,8)] - 2  #note these are in log10(w)

zoo_groups <- groups[which(groups$Type == "Zooplankton"),] #just keep the zooplankton

input <- enviro#[ID,]

# stable_zoomizer <- readRDS(paste0("Output/", zoomssjobname, "_ZooMSS_", ID_char,".RDS"))
stable_zoomizer <- readRDS("Output/20220519_15deg_chlovarying_ZooMSS_0013.RDS")


times <- length(getTimes(stable_zoomizer))
stable_zoo <- colMeans(stable_zoomizer@n[ceiling(times/2+1):times,,]) 

# stable_zoomizer@n <- sweep(stable_zoomizer@n, 2, 2.5 * stable_zoomizer@params@species_params$Carbon, "*") #convert to carbon
intercept <- mean(getCommunitySlope(stable_zoomizer,
                                    species = which(groups$Type == "Zooplankton"),
                                    max_w = 1e-5,
)$intercept[ceiling(times/2+1):times])
slope <- mean(getCommunitySlope(stable_zoomizer,
                                species = which(groups$Type == "Zooplankton"),
                                max_w = 1e-5,
)$slope[ceiling(times/2+1):times])


new <- stable_zoo[which(groups$Type == "Zooplankton"),]

temp_eff <- 2.^((input$sst - 30)/10)
```

Now set up the fish part of the model

```{r}
fish_params <- newTraitParams(no_sp = 13,
                                min_w = 10^(-3),
                                min_w_inf = 10^1,
                                max_w_inf = 10^7,
                                no_w = 10*(7-(-3))+1, # sets dx = 0.1
                                min_w_pp = 10^(-14.5),
                                alpha = 0.6, # * temp_eff, #takes care of assimilation efficiency & temp effect on Encounter
                                kappa = exp(intercept), #10^(input$phyto_int) * 10^-1,
                                lambda = 1 - slope, # 1 - input$phyto_slope,
                                gamma = 640 * temp_eff, #takes care of temp effect on PredRate and Encounter
                                # f0 = 0.6,
                                h = 40,
                                beta = 100,
                                sigma = 1.3,
                                # R_factor = 1.01, #RMax = RFactor * RDI, default 4. Note RDI 
                                reproduction_level = 0,
                                fc = 0,
                                ks = 0, #set metabolism to zero
                                ext_mort_prop = 0, #currently zeroed since this is fitted. No need for temp effect here, calculates from PredMort
                                knife_edge_size = 10  # min size for fishing
  )

# fish_params@species_params$gamma <- fish_params@species_params$gamma * temp_eff
# fish_params <- setSearchVolume(fish_params)

fish_params@other_params$temp_eff <- 2.^((input$sst - 30)/10)

```

Now setting up the zooplankton part of the model

```{r zoomodel}
zoo_params <- newZooMizerParams(groups = zoo_groups, input = input, fish_params = fish_params)

zoo_params@initial_n <- new[which(groups$Type == "Zooplankton"),]

zoo_params@species_params$interaction_resource[9] <- 1 # make jellies eat resource (i.e. fish)

## Zooplankton Mortality

#U-shaped mortality
M <- getExtMort(zoo_params) # pulls out senescence
z0pre <- rep(0, nrow(M))
z0pre[c(6,8)] <- 1 * (zoo_params@species_params$w_inf[c(6,8)])^(1/4) * fish_params@other_params$temp_eff #choose groups to apply to
allo_mort <- outer(z0pre, w(zoo_params)^(-1/4)) # mortality of z0pre * w^(-1/4)
zoo_params <- setExtMort(zoo_params, ext_mort = M + allo_mort)

# add fixed level of background mortality (so in total have mizer fixed background mortality + senescence)
# M <- getExtMort(zoo_params) # pulls out senescence
# z0pre <- rep(0, nrow(M))
# z0pre[c(6,8)] <- 0.6 #choose groups to apply to; default Mizer value 0.6
# mu_b <- z0pre * zoo_params@species_params$w_inf^(-1/4) * fish_params@other_params$temp_eff
# zoo_params <- setExtMort(zoo_params, ext_mort = M + mu_b) # adds 0.6 * w_inf^(-1/4)

#Both
# M <- getExtMort(zoo_params) # pulls out senescence
# z0pre <- rep(0, nrow(M))
# z0pre[c(6,8)] <- 1 * (zoo_params@species_params$w_inf[c(6,8)])^(1/4) * fish_params@other_params$temp_eff #choose groups to apply to
# larv_mort <- outer(z0pre, w(zoo_params)^(-1/4)) # mortality of z0pre * w^(-1/4)
# 
# z0pre1 <- rep(0, nrow(M))
# z0pre1[c(6,8)] <- 6 * fish_params@other_params$temp_eff #choose groups to apply to; default Mizer value 0.6
# back_mort <- z0pre * zoo_params@species_params$w_inf^(-1/4)
# zoo_params <- setExtMort(zoo_params, ext_mort = M + larv_mort + back_mort) # adds 0.6 * w_inf^(-1/4)!

# #type II feeding for zoo
# zoo_params <- setRateFunction(zoo_params, "PredRate", "mizerPredRate")
# zoo_params <- setRateFunction(zoo_params, "FeedingLevel", "mizerFeedingLevel")
# zoo_params@species_params$h <- 40
# zoo_params <- setMaxIntakeRate(zoo_params)

#Density-dependent mortality
zoo_params@species_params$mu0DD <- rep(80, 9) * zoo_params@species_params$w_inf
zoo_params@species_params$mu0DD[6] <- zoo_params@species_params$mu0DD[6] * 50 #turn up density-dependent mortality on krill
zoo_params <- setRateFunction(zoo_params, "Mort", "totalMortDD")
```

Combine the models by registering the zooplankton model as a component of the fish model and the resource, and set the initial conditions.

```{r combinemodels}
fish_params <- fish_params %>%
  setComponent(component = "zoo",
               initial_value = zoo_params@initial_n,
               dynamics_fun = "zoo_dynamics",
               mort_fun = "zoo_predation",
               component_params = list(params = zoo_params)
  ) %>%
  setResource(resource_dynamics = "resource_zooMizer")



# this line not needed if gamma is set to be temperature-dependent
# fish_params <- setRateFunction(fish_params, "PredRate", "PredRate_temp")



initialNResource(fish_params) <- resource_zooMizer(fish_params, fish_params@initial_n_other) #TODO: set this to be stable state of a regular ZooMizer run
# plotSpectra(fish_params, wlim = c(10^-14.5, NA), total = FALSE)

# initialN(fish_params) <- get_initial_n(fish_params) * 10^3 # adjust initial n for fish
initialNResource(fish_params@other_params$zoo$params) <- initialNResource(fish_params@other_params$zoo$params)
fish_idx <- (length(fish_params@w_full)-length(fish_params@w) + 1):length(fish_params@w_full)
initialNResource(fish_params@other_params$zoo$params)[fish_idx] <- initialNResource(fish_params@other_params$zoo$params)[fish_idx] + colSums(initialN(fish_params))
# fish_params <- setParams(fish_params)

# fish_params@species_params$R_max <- readRDS("data/rmaxs.RDS")/10
# fish_params@species_params$erepro <- 1
# fish_params@species_params$R_max <- fish_params@resource_params$kappa * fish_params@species_params$w_inf ^(-1.5)
# fish_params@species_params$R_max <- fish_params@species_params$R_max * 10^(5+3*log10(input$chlo))
# fish_params <- setBevertonHolt(fish_params, erepro = 1)
# fish_params@species_params$R_max <- fish_params@species_params$R_max * 10000
# fish_params <- setRateFunction(fish_params, "FeedingLevel", "FeedingLevel_type3")
# fish_params <- setExtMort(fish_params, ext_mort = array(data=0.1, dim=dim(fish_params@mu_b)))
```

## Run the coupled model

Finally, we run the model!

```{r runzoomizer, eval = FALSE}
out <- project(fish_params, t_max = 1000, dt = 0.01, effort = 0)

saveRDS(out, file = paste0("Output/", jobname, "_ZooMizer.RDS"))

```

```{r existingdata, include=FALSE}
out <- readRDS("Output/20220531_15deg_chlovarying_reprolevel0_alphapt6_1000yr_ZooMizer_0013.RDS")
```

## Look at results

We'll start with a summary plot:

```{r summaryplot}

source("ZooMizerPlots.R")
source("ZooMizerSummaryFunctions.R")
library(patchwork)
summary_plot(out)

```

Do the growth rates agree with data?

```{r growthsplot}

source("ZooMizerSummaryFunctions.R")
source("ZooMizerResourceFunctions.R")
source("GrowthPlots.R")

plotZooMizerGrowthsComparison(out, sst = 15)

```

