---
title: "Setting up a zooplankton model using mizer"
author: Patrick Sykes and Julia L. Blanchard
date: July 22, 2020
place: Brisbane
output:
  html_document: default
---


# Introduction

Here we will recreate the ZooMSS model (version 2) in Heneghan et al. (2020 in preparation) using mizer.

We begin with some setup of required packages.

```{r setup, message=FALSE, warning=FALSE}
#get required packages
library(devtools)
#most up to date master branch of mizer
#install_github("sizespectrum/mizer")
#install_github("astaaudzi/mizer-rewiring", ref = "temp-model-comp")
#documentation here:
#https://sizespectrum.org/mizer/dev/index.html
library(mizer)
require(tidyverse)

#remotes::install_github("sizespectrum/mizerExperimental")
library(mizerExperimental)

```

## Set-up mizer model

Next let's read in the parameters from ZooMSS.

```{r}

groups <-read_csv("data/TestGroups_mizer.csv")
groups$w_min <- 10^groups$w_min #convert from log10 values
groups$w_inf <- 10^groups$w_inf
groups$w_mat <- 10^groups$w_mat
groups$h <- 1e50 # should be Inf, but that breaks the calculations. Massive value still works out to effectively unlimited feeding as allowed in ZooMSS
groups$ks <- 0 #turn off standard metabolism
#todo - ramp up constant repro for coexistence

# read interaction matrix
# get the interaction matrix - actually I think we can leave this out. Default is all 1s, which is the same as in ZooMSS. Included for completeness; it may be useful in future to keep this in.
theta <- readRDS("data/zoomss_inter.rds")[,-1]

 
```

We will pass these parameters to mizer to set up a new multispecies model.

TODO: adjust parameters here.

```{r}


ID <- 223 #index of environmental data to choose
envirofull <- readRDS("data/envirofull_20200317.RDS")
enviro_row <- envirofull[envirofull$cellID==ID,]


#set up the fixed phyoplankton spectrum
phyto_fixed <- function(params, n, n_pp, n_other, rates, dt = 0.01, kappa = 10^enviro_row$phyto_int, lambda = 1-enviro_row$phyto_slope,...) {
    n_pp <- kappa*params@w_full^(-lambda) #returns the fixed spectrum at every time step
    n_pp[params@w_full>params@resource_params$w_pp_cutoff] <- 0
    return(n_pp)
}

mf.params <- newMultispeciesParams(species_params=groups,
                                   interaction=NULL, #NULL sets all to 1, no strict herbivores
                                   no_w = 178, #number of zoo+fish size classes;
                                   min_w_pp = 10^(-14.4), #minimum phyto size. Note: use -14.4, not -14.5, otherwise it makes an extra size class
                                   w_pp_cutoff = 10^(enviro_row$phyto_max), #maximum phyto size
                                   n = 0.7, #The allometric growth exponent used in ZooMSS
                                   z0pre = 1, #external mortality (senescence)
                                   z0exp = 0.3,
                                   resource_dynamics = "phyto_fixed",
                                   kappa = 10^(enviro_row$phyto_int), 
                                   lambda = 1-enviro_row$phyto_slope,
                                   RDD = constantRDD(species_params = groups) #first go at this
                                   #pred_kernel = ... #probably easiest to just import this/pre-calculate it, once dimensions are worked out
)

#checks
length(which(mf.params@initial_n_pp>0)) == length(seq(-14.5,enviro_row$phyto_max, by = 0.1))


```


Now do some fiddling to make the new MizerParams object match the ZooMSS parameters. Note: this chunk is adapted from `fZooMSS_setup.R`, found at https://github.com/MathMarEcol/ZooMSS/.

```{r}
setZooMizerConstants <- function(params, Groups, sst){
  #### CALCULATES CONSTANT BITS OF THE MODEL FUNCTIONS FOR EACH GROUP
  SearchVol <- getSearchVolume(params)
  M_sb <- getExtMort(params)
  ZSpre <- 1 # senescence mortality prefactor
  ZSexp <- 0.3 # senescence mortality exponent
  
  pred_kernel <- getPredKernel(params)
  prey_weight_matrix <- matrix(params@w_full, nrow = length(params@w), ncol = length(params@w_full), byrow = TRUE)
  pred_weight_matrix <- matrix(params@w, nrow = length(params@w), ncol = length(params@w_full))
  
    for(i in 1:nrow(params@species_params)){
    ## Senescence mortality
    if(params@species_params$Type[i] == "Zooplankton"){
      M_sb[i,] <- ZSpre*(params@w/(params@species_params$w_mat[i]))^ZSexp
      M_sb[i, params@species_params$w_max[i] < params@w] <- 0
      M_sb[i, params@species_params$w_mat[i] > params@w] <- 0
    }
    
    if(params@species_params$Type[i] == "Fish"){
      M_sb[i,] <- 0.1*ZSpre*(params@w/(params@species_params$w_mat[i]))^ZSexp
      M_sb[i, params@species_params$w_max[i] < params@w] <- 0
      M_sb[i, params@species_params$w_mat[i] > params@w] <- 0
    }
    
    ### Search volume
    SearchVol[i,] <- (params@species_params$gamma[i])*(params@w^(params@species_params$q[i]))
    SearchVol[i, params@species_params$w_max[i] < params@w] <- 0
    SearchVol[i, params@species_params$w_min[i] > params@w] <- 0
    
    ### Predation Kernels
    if(is.na(params@species_params$PPMRscale[i]) == FALSE){ # If group has an m-value (zooplankton)
      # Calculate PPMR for zooplankton, which changes according to body-size (Wirtz, 2012)
      D.z <- 2*(3*params@w*1e12/(4*pi))^(1/3) # convert body mass g to ESD (um)
      betas <- (exp(0.02*log(D.z)^2 - params@species_params$PPMRscale[i] + 1.832))^3 # Wirtz's equation
      beta_mat <- matrix(betas, nrow = length(params@w), ncol = length(params@w_full))
      
      # Calculate feeding kernels
      pred_kernel[i, , ] <- exp(-0.5*(log((beta_mat*prey_weight_matrix)/
                                            pred_weight_matrix)/params@species_params$FeedWidth[i])^2)/
        sqrt(2*pi*params@species_params$FeedWidth[i]^2)
      # The feeding kernel of filter feeders is not expected to change  with increasing size so we fix it here
      
      # if (param$fixed_filterPPMR == TRUE){
      if(i == 3){
        pred_kernel[i, , ] <- matrix(pred_kernel[i,44,], nrow = length(params@w), ncol = length(params@w_full), byrow = TRUE)
      }
      if(i == 8){
        pred_kernel[i, , ] <- matrix(pred_kernel[i,61,], nrow = length(params@w), ncol = length(params@w_full), byrow = TRUE)
      }
      # }
      
    } else { # If group does not have an m-value (fish)
      beta_mat <- matrix(params@species_params$PPMR[i], nrow = length(params@w), ncol = length(params@w_full))
      
      # Calculate feeding kernels
      pred_kernel[i, , ] <- exp(-0.5*(log((beta_mat*prey_weight_matrix)/
                                            pred_weight_matrix)/params@species_params$FeedWidth[i])^2)/
        sqrt(2*pi*params@species_params$FeedWidth[i]^2)
    }
    
  }
  
  #temperature effect 
  temp_eff <-  matrix(2.^((sst - 30)/10), nrow = length(params@species_params$species), ncol = length(params@w))
  M_sb <- temp_eff * M_sb # Incorporate temp effect on senscence mortality
  
  
  params@initial_n_pp <- params@resource_params$kappa * params@w_full^(-params@resource_params$lambda)
  params@initial_n_pp[params@w_full > params@resource_params$w_pp_cutoff] <- 0
  
  
  a_dynam <- (params@resource_params$kappa)*(params@w[1]^(2-params@resource_params$lambda)) # calculate coefficient for initial dynamic spectrum, so that N(w_phyto) equals N(w_dynam) at w[1]
  
  # Initial abundances form a continuation of the plankton spectrum, with an abundance slope of -1 (normalised abundance slop of -2)
  tempN <- matrix(a_dynam*(params@w)^-2, nrow = nrow(params@species_params), ncol = length(params@w), byrow = TRUE)
  props_z <- params@species_params$Prop[params@species_params$Type=="Zooplankton"] # Zooplankton proportions
  tempN[params@species_params$Type=="Zooplankton",] <- props_z * tempN[params@species_params$Type=="Zooplankton",] # Set abundances of diff zoo groups based on smallest size class proportions
  tempN[params@species_params$Type=="Fish",] <- (1/sum(params@species_params$Type=="Fish")) * tempN[params@species_params$Type=="Fish",] # Set abundandances of fish groups based on smallest size class proportions
  
  # For each group, set densities at w > Winf and w < Wmin to 0
  tempN[unlist(tapply(round(log10(params@w), digits = 2), 1:length(params@w), function(wx,Winf) Winf < wx, Winf = log10(params@species_params$w_inf)))] <- 0
  tempN[unlist(tapply(round(log10(params@w), digits = 2), 1:length(params@w), function(wx,Wmin) Wmin > wx, Wmin = log10(params@species_params$w_min)))] <- 0
  dimnames(tempN) <- dimnames(params@initial_n)
  params@initial_n <- tempN
  
  params <- setExtMort(params, z0 = M_sb)
  params <- setSearchVolume(params, SearchVol)
  params <- setPredKernel(params, pred_kernel)
  
  return(params)
}


mf.params <- setZooMizerConstants(params = mf.params, Groups = groups, sst= enviro_row$sst)


```

Try running it:

```{r}

mf.params <- setParams(mf.params)

sim <- project(mf.params)
plot(sim) #note feeding level means satiation - 0 since there's no satiation in this model.
plotlyBiomass(sim)
#plotlyGrowthCurves(sim,species="macrozooplankton")
#plotlyFeedingLevel(sim)
# feeding level satiation for some groups, except for the seabirds
# macrozooplankton - they are not growing enough,why?
#tuneParams(mf.params)
#plotlyGrowthCurves(sim,percentage = T)
plotlySpectra(sim, power = 1)
```


Next thing to do is reproduction. In ZooMSS, this is handled by simply setting the abundance in the smallest size class to be a fixed proportion of the community size spectrum; in short

$$ N_i(w_{min}(i)) = prop(i) \sum_{j \neq i} N_j(w_{min}(i)),$$
where $N_i(w)$ is the density of species $i$ in weight class $w$, and $prop(i)$ is a (fixed) proportion depending on the species.

In mizer, reproduction is linked to metabolism. The abundance in the smallest size class is proportional to the energy available to mature individuals for reproduction - i.e. the energy left over after subtracting metabolic costs (including energy to support growth) from the energy assimilated (by mature individuals) from feeding on prey.

Now, to recreate this in mizer, we need to rewrite mizer's `project_simple()` function. We do this by making a new function, `new_project_simple()`, and using it in place of the default one:

TODO: insert remaining temperature effects in here too.

```{r}
new_project_simple <- function(params, n, n_pp, n_other, t, dt, steps, 
                           effort, resource_dynamics_fn, other_dynamics_fns,
                           rates_fns, ...) {    
    # Handy things
    no_sp <- nrow(params@species_params) # number of species
    no_w <- length(params@w) # number of fish size bins
    idx <- 2:no_w
    w_max_idx <- params@w_min_idx
    for (i in 1:length(w_max_idx)) {
     w_max_idx[i] <- which(round(log10(params@w),2)==round(log10(params@species_params$w_inf[i]),2)) 
    }
    
    # Hacky shortcut to access the correct element of a 2D array using 1D notation
    # This references the egg size bracket for all species, so for example
    # n[w_min_idx_array_ref] = n[,w_min_idx]
    w_min_idx_array_ref <- (params@w_min_idx - 1) * no_sp + (1:no_sp)
    # Matrices for solver
    a <- matrix(0, nrow = no_sp, ncol = no_w)
    b <- matrix(0, nrow = no_sp, ncol = no_w)
    S <- matrix(0, nrow = no_sp, ncol = no_w)
    
    for (i_time in 1:steps) {
        r <- rates_fns$Rates(
            params, n = n, n_pp = n_pp, n_other = n_other,
            t = t, effort = effort, rates_fns = rates_fns, ...)
        
        # Update time
        t <- t + dt
        
        # Update other components
        n_other_current <- n_other  # So that the resource dynamics can still 
        # use the current value
        for (component in names(params@other_dynamics)) {
            n_other[[component]] <-
                other_dynamics_fns[[component]](
                    params,
                    n = n,
                    n_pp = n_pp,
                    n_other = n_other_current,
                    rates = r,
                    t = t,
                    dt = dt,
                    component = component,
                    ...
                )
        }
        
        # Update resource
        n_pp <- resource_dynamics_fn(params, n = n, n_pp = n_pp,
                                     n_other = n_other_current, rates = r,
                                     t = t, dt = dt, ...)
        
        # Iterate species one time step forward:
        # a_{ij} = - g_i(w_{j-1}) / dw_j dt
        a[, idx] <- sweep(-r$e_growth[, idx - 1, drop = FALSE] * dt, 2,
                          params@dw[idx], "/")
        # b_{ij} = 1 + g_i(w_j) / dw_j dt + \mu_i(w_j) dt
        b[] <- 1 + sweep(r$e_growth * dt, 2, params@dw, "/") + r$mort * dt
        # S_{ij} <- N_i(w_j)
        S[,idx] <- n[, idx, drop = FALSE]
        # Update n
        # for (i in 1:no_sp) # number of species assumed small, so no need to 
        #                      vectorize this loop over species
        #     for (j in (params@w_min_idx[i]+1):no_w)
        #         n[i,j] <- (S[i,j] - A[i,j]*n[i,j-1]) / B[i,j]
        # This is implemented via Rcpp
        n <- inner_project_loop(no_sp = no_sp, no_w = no_w, n = n,
                                A = a, B = b, S = S,
                                w_min_idx = params@w_min_idx)
    }
        
        # Update first and last size groups of n
        #TODO: Make this a little less hacky
        n[1,1] <- params@species_params$Prop[1]*n_pp[length(params@w_full)-length(params@w)+1]
        n[1,w_max_idx[1]] <- 0
        for (i in 2:no_sp) {
          n[i, w_max_idx[i]] <- 0
          if(params@species_params$Type[i] != "Fish"){
            n[i,params@w_min_idx[i]] <- params@species_params$Prop[i] * sum(n[-i,params@w_min_idx[i]])
            }
          if(params@species_params$Type[i] == "Fish"){
            n[i,params@w_min_idx[i]] <- 1/3 * sum(n[1:9,params@w_min_idx[i]])
            }
        }
            
    return(list(n = n, n_pp = n_pp, n_other = n_other, rates = r))
}


#assign new project function in namespace
environment(new_project_simple) <- asNamespace('mizer')
assignInNamespace("project_simple", new_project_simple, ns = "mizer")
```


Now let's try that out:

```{r}
sim2 <- project(mf.params, t_max = 100, dt = 001)
plot(sim2)
plotlySpectra(sim2, power = 0)
```

Still missing is the differential prey nutrition based on prey carbon content. We'll do that by editing the `mizerEncounter()` function found in `project_methods.R`, and using `setRateFunction(params, "Encounter", "myEncounter")`. It might be possible to instead edit the chunk above too to incorporate the different method (since we've gone there anyway...)

```{r}

#set up matrix of pred nutrition given prey, dims (pred species) x (prey species)
assim_eff = matrix(mf.params@species_params$GrossGEscale * mf.params@species_params$Carbon, nrow = nrow(mf.params@species_params), ncol = nrow(mf.params@species_params))
get_filterfeeders <- which(mf.params@species_params$FeedType == "FilterFeeder")

cc_phyto <- 0.1 #carbon content of phytoplankton
assim_phyto <- mf.params@species_params$GrossGEscale * cc_phyto #assimilation efficiency when eating phytoplankton

#TODO: make the assim_eff a parameter in the MizerParams object
for (i in get_filterfeeders) {
  assim_eff[,i] <- assim_eff[,i] / mf.params@species_params$Carbon[i]
}

new_Encounter <- function(params, n, n_pp, n_other, t, ...) {

    # idx_sp are the index values of params@w_full such that
    # params@w_full[idx_sp] = params@w
    idx_sp <- (length(params@w_full) - length(params@w) + 1):length(params@w_full)
    
    # Note: removed the FFT code because it does not apply to this case.
    
    # If the feeding kernel does not have a fixed predator/prey mass ratio
    # then the integral is not a convolution integral and we can not use fft.
    # In this case we use the code from mizer version 0.3
    
        # n_eff_prey is the total prey abundance by size exposed to each
        # predator (prey not broken into species - here we are just working out
        # how much a predator eats - not which species are being eaten - that is
        # in the mortality calculation
        # \sum_j \theta_{ij} N_j(w_p) w_p dw_p
        assim_prey <- assim_eff * params@interaction
        n_eff_prey <- sweep(assim_prey %*% n, 2, 
                            params@w * params@dw, "*", check.margin = FALSE) 
        # pred_kernel is predator species x predator size x prey size
        # So multiply 3rd dimension of pred_kernel by the prey biomass density
        # Then sum over 3rd dimension to get consumption rate of each predator by 
        # predator size
        # This line is a bottle neck
        phi_prey_species <- rowSums(sweep(
            params@pred_kernel[, , idx_sp, drop = FALSE],
            c(1, 3), n_eff_prey, "*", check.margin = FALSE), dims = 2)
        # Eating the background
        # This line is a bottle neck
        phi_prey_background <- assim_phyto * params@species_params$interaction_resource *
            rowSums(sweep(
            params@pred_kernel, 3, params@dw_full * params@w_full * n_pp,
            "*", check.margin = FALSE), dims = 2)
        encounter <- params@search_vol * (phi_prey_species + phi_prey_background)
    
    dimnames(encounter) <- dimnames(params@metab)
    
    # Add contributions from other components
    for (i in seq_along(params@other_encounter)) {
        encounter <- encounter + 
            do.call(params@other_encounter[[i]], 
                    list(params = params,
                         n = n, n_pp = n_pp, n_other = n_other,
                         component = names(params@other_encounter)[[i]], ...))
    }
    
    return(encounter)
}
    
mf.params2 <- setRateFunction(mf.params, "Encounter", "new_Encounter")

sim3 <- project(mf.params2)
plot(sim3)
plotlySpectra(sim3, power=0)
```

Looking good, but we still need to include the temperature effects on the other processes.

```{r}

temp_eff <-  matrix(2.^((enviro_row$sst - 30)/10), nrow = length(mf.params@species_params$species), ncol = length(mf.params@w))

new_PredRate <- function (params, n, n_pp, n_other, t, feeding_level, ...) 
{
    no_sp <- dim(params@interaction)[1]
    no_w <- length(params@w)
    no_w_full <- length(params@w_full)
    if (length(params@ft_pred_kernel_p) == 1) {
        n_total_in_size_bins <- sweep(n, 2, params@dw, "*", 
            check.margin = FALSE)
        pred_rate <- sweep(params@pred_kernel, c(1, 2), (1 - 
            feeding_level) * params@search_vol * n_total_in_size_bins * temp_eff, 
            "*", check.margin = FALSE)
        pred_rate <- colSums(aperm(pred_rate, c(2, 1, 3)), dims = 1)
        return(pred_rate)
    }
    idx_sp <- (no_w_full - no_w + 1):no_w_full
    Q <- matrix(0, nrow = no_sp, ncol = no_w_full)
    Q[, idx_sp] <- sweep((1 - feeding_level) * params@search_vol * temp_eff *
        n, 2, params@dw, "*")
    pred_rate <- Re(t(mvfft(t(params@ft_pred_kernel_p) * mvfft(t(Q)), 
        inverse = TRUE)))/no_w_full
    return(pred_rate * params@ft_mask)
}

new_EReproAndGrowth <- function (params, n, n_pp, n_other, t, encounter, feeding_level, 
    ...) 
{
    sweep((1 - feeding_level) * encounter * temp_eff, 1, params@species_params$alpha, 
        "*", check.margin = FALSE) - params@metab
}

mf.params3 <- setRateFunction(mf.params2, "PredRate", "new_PredRate")
mf.params3 <- setRateFunction(mf.params3, "EReproAndGrowth", "new_EReproAndGrowth")

#fix up energy to reproduction
mf.params3 <- setReproduction(mf.params3, repro_prop = matrix(0, nrow = nrow(mf.params3@psi), ncol = ncol(mf.params3@psi)))



sim4 <- project(mf.params3, t_max = 100, dt = 0.01)
plot(sim4)
plotlySpectra(sim4, power = 0)

```


Something is not quite right...Here's a very hacky way to get the comparison...I've run the model to 1000 years with `dt=0.01` for 20 different combinations of SST and chlorophyll-a and matched them to the relevant runs of ZooMSS (without diffusion).

```{r}

require(ggplot2)

hack <- sim2 #making a "fake" MizerSim object that we can use plots on.

#match ZooMSS data to that collected from zoomizer
enviro <- readRDS("data/enviro_test20.RDS")
zoomssgrid <- readRDS("Output/res_20210309_20cells.RDS")
zoomssgrid <- zoomssgrid[rank(enviro$CellID)]

w <- matrix(hack@params@w, nrow = nrow(hack@params@species_params), ncol = length(hack@params@w), byrow = TRUE)
dw <- matrix(hack@params@dw, nrow = nrow(hack@params@species_params), ncol = length(hack@params@dw), byrow = TRUE)

#load in ZooMizer data for comparison
zoomizergrid <- readRDS("test_grid_20210317.RDS")

tmax <- dim(zoomizergrid[[1]]@n)[1]

pzoomizer <- list()
for (i in 1:length(zoomizergrid)) {
  zoomizergrid[[i]]@n[tmax,,] <- #apply(as.array(zoomizergrid[[i]]@n[ceiling(tmax/2):tmax,,]),c(2,3),'mean')
    colMeans(zoomizergrid[[i]]@n[(ceiling(0.5*(dim(zoomizergrid[[i]]@n)[1])):dim(zoomizergrid[[i]]@n)[1]),,], dims = 1)#apples-to-apples comparison
  pzoomizer[[i]] <- plotSpectra(zoomizergrid[[i]], time_range = tmax-1, power = 1)+labs(title = paste("ZooMizer plot",i))
}

#fill fake MizerSim object with ZooMSS data and plot
pzoomss <- list()
for (i in 1:length(zoomssgrid)) {
   hack@n[i,,] <- zoomssgrid[[i]]/ dw
   hack@n_pp[i,] <- zoomizergrid[[i]]@n_pp[1,]
   pzoomss[[i]] <- plotSpectra(hack, time_range = i-1, power = 1)+labs(title = paste("ZooMSS plot",i))
}





#have a look
pzoomss[[ceiling(length(zoomizergrid))/2]]
pzoomizer[[ceiling(length(zoomizergrid))/2]]

#compare them all
# p <- list(length(zoomizergrid)*2)
# for (i in 1:length(zoomizergrid)) {
#     p[[2*i-1]] <- pzoomss[[i]]
#     p[[2*i]] <- pzoomizer[[i]]
# }
#p

testdf <- data.frame(zoomsstcb = numeric(length(zoomizergrid)),zoomizertcb = numeric(length(zoomizergrid)))
for (i in 1:nrow(testdf)) {
testdf$zoomsstcb[i] = sum(hack@n[i,,]*w*dw)
testdf$zoomizertcb[i] = sum(zoomizergrid[[i]]@n[tmax,,]*w*dw)
}

(scatter <- ggplot(testdf, aes(x=zoomsstcb, y=zoomizertcb))+
    geom_abline(slope = 1, intercept = 0) +
    geom_point()+
    geom_smooth(method = 'lm'))

ggsave("zoomss_zoomizer_comparison_tcb.png",scatter)

testdf$chlo <- enviro$chlo[1:nrow(testdf)]
(chlotcb <- ggplot(data=testdf, aes(x=log10(chlo),y=log10(zoomizertcb)))+
  geom_point())
ggsave("zoomss_tcb_vs_chlo.png", chlotcb)

fit <- lm(testdf$zoomizertcb~testdf$zoomsstcb)


```


