---
title: "Exploring ZooMizer variation across chlorophyll concentration"
author: "Patrick Sykes"
date: "18/05/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we will look at how chlorophyll affects growth rates and relative biomass in ZooMizer.

## Load packages

We will be using Mizer functions and some tidyverse, so let's load those

```{r packages}
library(mizer)
library(tidyverse)
library(assertthat)
library(foreach)

```

## Run the code

Give the job a name and set up the environment data:

```{r}

jobname <- "20220525_15deg_chlovarying_NSrepro1000yr"

enviro <- data.frame(sst = 15, chlo = 10^seq(-1.5, 1.5, by=0.125))
source("fZooMSS_CalculatePhytoParam.R")
enviro <- fZooMSS_CalculatePhytoParam(enviro)
enviro$tmax <- 1000

```

I'm going to do this in parallel (with `foreach` and `%doPar%`), so I set up the virtual cluster first

```{r cluster, echo=FALSE}
library(doParallel)
cl <- makePSOCKcluster(max(1, detectCores()-1), outfile="")     ## set up cores-1 machines
registerDoParallel(cl, cores = (max(1, detectCores()-1)))
clusterEvalQ(cl, lapply(c("mizer", "assertthat", "tidyverse"), require, character.only = TRUE)) %>% invisible()
```

Then I can run the code over the cluster

```{r runzoomizer, eval=FALSE}

walltimes <- foreach(ID = 1:25,
                .combine = rbind,     
                .packages = c("mizer", "assertthat", "tidyverse"),
                .export = c("jobname", "enviro")
) %dopar% {
  
start_time <- Sys.time()

#job specifics
Groups <- read.csv("data/TestGroups_mizer.csv") # Load in functional group information

zoomssjobname <- "20220519_15deg_chlovarying" #job name used on queue
# jobname <- '20220212_type1'

# ID <- as.integer(Sys.getenv('PBS_ARRAY_INDEX'))
ID_char <- sprintf("%04d",ID)

# Choose environmental data to use
  enviro <- data.frame(sst = 15, chlo = 10^seq(-1.5, 1.5, 0.125))
  source("fZooMSS_CalculatePhytoParam.R")
  enviro <- fZooMSS_CalculatePhytoParam(enviro)
  enviro$tmax <- 1000
  enviro$dt <- 0.01

source("uncoupledmodel.R")
source("ZooMizerResourceFunctions.R")

# environment(new_project_simple) <- asNamespace('mizer')
# assignInNamespace("project_simple", new_project_simple, ns = "mizer")
# 
# # Run the ZooMSS model inside Mizer:
# phyto_fixed <- function(params, n, n_pp, n_other, rates, dt, kappa = 10^enviro$phyto_int[i], lambda= 1-enviro$phyto_slope[i], ... ) {
#   npp <- kappa*params@w_full^(-lambda) #returns the fixed spectrum at every time step
#   npp[params@w_full > params@resource_params$w_pp_cutoff] <- 0
#   return(npp)
# }
# 
# # density-dependent mortality
# # totalMortDD <- function(params, n, n_pp, n_other, t, f_mort, pred_mort, ...){
# #   mort <- pred_mort + params@mu_b + f_mort
# #   # Add contributions from other components
# #   # for (i in seq_along(params@other_mort)) {
# #   #   mort <- mort + 
# #   #     do.call(params@other_mort[[i]], 
# #   #             list(params = params,
# #   #                  n = n, n_pp = n_pp, n_other = n_other, t = t,
# #   #                  component = names(params@other_mort)[[i]], ...))
# #   # }
# #   ddmort_raw <- sweep(n, "w", params@w^params@species_params$q[1] * params@dw, "*")
# #   ddmort <- ddmort_raw * 0
# #   sp_sel <- 1:nrow(ddmort) #c(6,7,8,9)
# #   mu0 <- 80 # Benoit & Rochet value = 80, but no density-independent mortality in that model
# #   ddmort[sp_sel,] <- mu0 * ddmort_raw[sp_sel,] #* params@species_params$w_inf
# #   ddmort[6,] <- ddmort[6,] * 50
# #   return(mort + ddmort)
# # }
# 
# zoomss <- fZooMizer_run(groups = Groups, input = enviro[ID,], no_w = (1*177+1)) # dx=0.1
# saveRDS(zoomss, file = paste0("Output/", zoomssjobname, "_ZooMSS_", ID_char,".RDS"))

mid <- Sys.time()
# 
environment(project_simple) <- asNamespace('mizer')
assignInNamespace("project_simple", project_simple, ns = "mizer")

ZooMizer_coupled <- function(ID, tmax = 1000, effort = 0) {
  groups <- read.csv("data/TestGroups_mizer.csv") #load in ZooMSS groups
  
  #lower "starting proportion" of salps and chaetognaths
  # groups$Prop[c(7,8)] <- 0.1*groups$Prop[c(7,8)]
  # #lower "starting proportion" of euphausiids
  # groups$Prop[6] <- 0.1*groups$Prop[c(6)]
  
  # change start of senescence for salps and euphs
  # groups$w_mat[c(6,8)] <- groups$w_mat[c(6,8)] - 2  #note these are in log10(w)
  
  zoo_groups <- groups[which(groups$Type == "Zooplankton"),] #just keep the zooplankton
  
  enviro <- data.frame(sst = 15, chlo = 10^seq(-1.5, 1.5, by=0.125))
  source("fZooMSS_CalculatePhytoParam.R")
  enviro <- fZooMSS_CalculatePhytoParam(enviro)
  enviro$tmax <- 1000
  
  input <- enviro[ID,]
  
  stable_zoomizer <- readRDS(paste0("Output/", zoomssjobname, "_ZooMSS_", ID_char,".RDS"))
  
  times <- length(getTimes(stable_zoomizer))
  stable_zoo <- colMeans(stable_zoomizer@n[ceiling(times/2+1):times,,]) 
  
  # stable_zoomizer@n <- sweep(stable_zoomizer@n, 2, 2.5 * stable_zoomizer@params@species_params$Carbon, "*") #convert to carbon
  intercept <- mean(getCommunitySlope(stable_zoomizer,
                                      species = which(groups$Type == "Zooplankton"),
                                      max_w = 1e-5,
  )$intercept[ceiling(times/2+1):times])
  slope <- mean(getCommunitySlope(stable_zoomizer,
                                  species = which(groups$Type == "Zooplankton"),
                                  max_w = 1e-5,
  )$slope[ceiling(times/2+1):times])
  
  
  new <- stable_zoo[which(groups$Type == "Zooplankton"),]
  
  temp_eff <- 2.^((input$sst - 30)/10)
  R_factor = 4
  fish_params <- newTraitParams(no_sp = 13,
                                min_w = 10^(-3),
                                min_w_inf = 10^1,
                                max_w_inf = 10^7,
                                no_w = 10*(7-(-3))+1, # sets dx = 0.1
                                min_w_pp = 10^(-14.5),
                                alpha = 0.25, # * temp_eff, #takes care of assimilation efficiency & temp effect on Encounter
                                kappa = exp(intercept), #10^(input$phyto_int) * 10^-1,
                                lambda = 1 - slope, # 1 - input$phyto_slope,
                                gamma = 640 * temp_eff, #takes care of temp effect on PredRate and Encounter
                                # f0 = 0.6,
                                h = 40,
                                beta = 100,
                                sigma = 1.3,
                                # R_factor = 1.01, #RMax = RFactor * RDI, default 4. Note RDI 
                                reproduction_level = 1/R_factor,
                                fc = 0,
                                ks = 0, #set metabolism to zero
                                ext_mort_prop = 0, #currently zeroed since this is fitted. No need for temp effect here, calculates from PredMort
                                knife_edge_size = 10  # min size for fishing
  )
  
  # fish_params@species_params$gamma <- fish_params@species_params$gamma * temp_eff
  # fish_params <- setSearchVolume(fish_params)
  
  zoo_params <- newZooMizerParams(groups = zoo_groups, input = input, fish_params = fish_params)
  
  zoo_params@initial_n <- new[which(groups$Type == "Zooplankton"),]
  
  zoo_params@species_params$interaction_resource[9] <- 1 # make jellies eat resource (i.e. fish)
  
  ## Zooplankton Mortality
  fish_params@other_params$temp_eff <- 2.^((input$sst - 30)/10)
  
  
  #U-shaped mortality
  M <- getExtMort(zoo_params) # pulls out senescence
  z0pre <- rep(0, nrow(M))
  z0pre[c(6,8)] <- 1 * (zoo_params@species_params$w_inf[c(6,8)])^(1/4) * fish_params@other_params$temp_eff #choose groups to apply to
  allo_mort <- outer(z0pre, w(zoo_params)^(-1/4)) # mortality of z0pre * w^(-1/4)
  zoo_params <- setExtMort(zoo_params, ext_mort = M + allo_mort)
  
  # hockey stick mortality (mizer background mortality + senescence)
  # M <- getExtMort(zoo_params) # pulls out senescence
  # z0pre <- rep(0, nrow(M))
  # z0pre[c(6,8)] <- 0.6 #choose groups to apply to; default Mizer value 0.6
  # mu_b <- z0pre * zoo_params@species_params$w_inf^(-1/4) * fish_params@other_params$temp_eff
  # zoo_params <- setExtMort(zoo_params, ext_mort = M + mu_b) # adds 0.6 * w_inf^(-1/4)
  
  #Both
  # M <- getExtMort(zoo_params) # pulls out senescence
  # z0pre <- rep(0, nrow(M))
  # z0pre[c(6,8)] <- 1 * (zoo_params@species_params$w_inf[c(6,8)])^(1/4) * fish_params@other_params$temp_eff #choose groups to apply to
  # larv_mort <- outer(z0pre, w(zoo_params)^(-1/4)) # mortality of z0pre * w^(-1/4)
  # 
  # z0pre1 <- rep(0, nrow(M))
  # z0pre1[c(6,8)] <- 6 * fish_params@other_params$temp_eff #choose groups to apply to; default Mizer value 0.6
  # back_mort <- z0pre * zoo_params@species_params$w_inf^(-1/4)
  # zoo_params <- setExtMort(zoo_params, ext_mort = M + larv_mort + back_mort) # adds 0.6 * w_inf^(-1/4)!
  
  # #type II feeding for zoo
  # zoo_params <- setRateFunction(zoo_params, "PredRate", "mizerPredRate")
  # zoo_params <- setRateFunction(zoo_params, "FeedingLevel", "mizerFeedingLevel")
  # zoo_params@species_params$h <- 40
  # zoo_params <- setMaxIntakeRate(zoo_params)
  
  zoo_params@species_params$mu0DD <- rep(80, 9) * zoo_params@species_params$w_inf
  zoo_params@species_params$mu0DD[6] <- zoo_params@species_params$mu0DD[6] * 50
  zoo_params <- setRateFunction(zoo_params, "Mort", "totalMortDD")
  
  fish_params <- fish_params %>%
    setComponent(component = "zoo",
                 initial_value = zoo_params@initial_n,
                 dynamics_fun = "zoo_dynamics",
                 mort_fun = "zoo_predation",
                 component_params = list(params = zoo_params)
    ) %>%
    setResource(resource_dynamics = "resource_zooMizer")
  
  
  
  # this line not needed if gamma is set to be temperature-dependent
  # fish_params <- setRateFunction(fish_params, "PredRate", "PredRate_temp")
  
  
  
  initialNResource(fish_params) <- resource_zooMizer(fish_params, fish_params@initial_n_other) #TODO: set this to be stable state of a regular ZooMizer run
  # plotSpectra(fish_params, wlim = c(10^-14.5, NA), total = FALSE)
  
  # initialN(fish_params) <- get_initial_n(fish_params) * 10^3 # adjust initial n for fish
  initialNResource(fish_params@other_params$zoo$params) <- initialNResource(fish_params@other_params$zoo$params)
  fish_idx <- (length(fish_params@w_full)-length(fish_params@w) + 1):length(fish_params@w_full)
  initialNResource(fish_params@other_params$zoo$params)[fish_idx] <- initialNResource(fish_params@other_params$zoo$params)[fish_idx] + colSums(initialN(fish_params))
  # fish_params <- setParams(fish_params)
  
  # fish_params@species_params$R_max <- readRDS("data/rmaxs.RDS")/10
  fish_params@species_params$erepro <- 1
  fish_params@species_params$R_max <- fish_params@resource_params$kappa * fish_params@species_params$w_inf ^(-1.5)
  # fish_params@species_params$R_max <- fish_params@species_params$R_max * 10^(5+3*log10(input$chlo))
  # fish_params <- setBevertonHolt(fish_params, erepro = 1)
  # fish_params@species_params$R_max <- fish_params@species_params$R_max * 10000
  # fish_params <- setRateFunction(fish_params, "FeedingLevel", "FeedingLevel_type3")
  # fish_params <- setExtMort(fish_params, ext_mort = array(data=0.1, dim=dim(fish_params@mu_b)))
  
  return(project(fish_params, t_max = tmax, dt = 0.01, effort = effort))
}  

out <- ZooMizer_coupled(ID, tmax = 1000, effort = 0)
saveRDS(out, file = paste0("Output/", jobname, "_ZooMizer_", ID_char,".RDS"))

end_time <- Sys.time()
c(ZooMSS = mid - start_time, ZooMizer = end_time - mid)
}

```

How long did those take to run?

```{r walltimes, eval=FALSE, include=FALSE}
walltimes
```

## Plots

OK, now prepare the run data for plotting:

```{r datalist}
sims <- foreach(ID = 1:25,
                .packages = c("mizer", "assertthat", "tidyverse")
                #.export = c("ZooMizer_coupled", "PredRate_temp")
) %dopar% {
  ID_char <- sprintf("%04d",ID)
  readRDS(paste0("Output/",jobname,"_ZooMizer_", ID_char,".RDS"))
}

```

### Growth plots

```{r growthplots}

source("ZooMizerSummaryFunctions.R")
source("ZooMizerResourceFunctions.R")
source("GrowthPlots.R")

plotZooMizerGrowthsComparison(sims[[1]])+ plot_annotation(title = paste0("SST = ", enviro[1,'sst'], ", chlo = ", round(enviro[1,'chlo'], 3)))#sst = 20, chlo = -1.5

plotZooMizerGrowthsComparison(sims[[11]])+ plot_annotation(title = paste0("SST = ", enviro[11,'sst'], ", chlo = ", round(enviro[13,'chlo'], 3)))#sst = 20, chlo = -0.5

plotZooMizerGrowthsComparison(sims[[21]])+ plot_annotation(title = paste0("SST = ", enviro[21,'sst'], ", chlo = ", round(enviro[25,'chlo'], 3)))#sst = 20, chlo = 0.5

```

### Biomass vs Chlorophyll

```{r biomchlo}

b <- foreach(ID=1:25, .combine = rbind) %dopar% {
  bioms <- getBiomass_ZooMizer(sims[[ID]], sims[[ID]]@params@other_params$zoo$params)
  rows <- ceiling(nrow(bioms)/2):nrow(bioms)
  colMeans(bioms[rows,])
}

df <- cbind(enviro[1:25,], b)
propdf <- df
propdf[,c(3:5,10:11)] <- propdf[,c(3:5,10:11)] / rowSums(propdf[,c(3:5,10:11)])
propdf[,12:18] <- propdf[,12:18] / rowSums(propdf[,12:18])
propdf[,19:31] <- propdf[,19:31] / rowSums(propdf[,19:31])

df2 <- gather(propdf, Group, Proportion, c(pico_biom:micro_biom,Flagellates:Ciliates))
gg1 <- ggplot(df2, aes(x=log10(chlo), y = Proportion, fill = Group)) + 
  geom_area() +
  labs(title = "Microzoo proportion")

df3 <- gather(propdf, Group, Proportion, Larvaceans:Jellyfish)
gg2 <- ggplot(df3, aes(x=log10(chlo), y = Proportion, fill = Group)) + 
  geom_area() +
  labs(title = "Zooplankton proportion")

df4 <- gather(propdf, Group, Proportion, "1":"13")
gg3 <- ggplot(df4, aes(x=log10(chlo), y = Proportion, fill = Group)) + 
  geom_area() +
  labs(title = "Fish proportion")

library(patchwork)
gg1
gg2
gg3

```

Slopes?

```{r slopes}

slopes <- foreach(ID=c(1:25), .combine = rbind) %dopar% { #cell 2 fails for some reason?
  colMeans(getCommunitySlope(sims[[ID]])[76:100,], na.rm = TRUE)
}


slopes <- cbind(enviro, slopes)

slopesplot <- ggplot(slopes, aes(x=log10(chlo)))+
  geom_point(aes(y=slope))
intsplot <- ggplot(slopes, aes(x=log10(chlo)))+
  geom_point(aes(y=intercept))

slopesplot / intsplot + plot_annotation(title = "Fish community slope + intercept")

phytoslopesplot <- ggplot(enviro, aes(x=log10(chlo), y = phyto_slope))+geom_point()
phytointsplot <- ggplot(enviro, aes(x=log10(chlo), y = phyto_int))+geom_point()

phytoslopesplot / phytointsplot + plot_annotation(title = "Phyto slope + intercept")

```
